How Beauty Trumps Performance: An Examination of Instructor Evaluations at the University of Texas

### 2023-04-20 outputs/paper/paper.qmd
reworked abstract to include findings + github link inlcluded

---
title: "Physical Appearance Affects Instructors Ratings More Than We Think"
author: "Marcin Jaczynski"
thanks: "Code and data are available at: __repo link HERE___"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "At the end of each semester of school, students are able to rate their professors and teachers through course evaluations. There are two schools of thought for course evaluations: they are repersentative of how well the teacher was able to teach a course, and if the students liked them or they are meaningless as only students who received a bad grade fill out the course evaluations and they are not representative of the teachers ability to teach. However, there are many factors that influece a teacher's rating. In this paper I observe how the beauty has an impact on their overall evaluations and see how other factors also play a role in the rating outcome."
format: pdf
bibliography: references.bib
---

@@

Teacher ratings have a direct impact on the validity of a teacher or professor. While there are a lot of factors that play on how students will perceive a teacher and what kind of rating they will give them at the end of the semester, one that is not taken into consideration many times is how beautiful they are. As such, I have chosen to try and answer "How does an instructors beauty, according to their students, affect the instructors course evaluation?" This paper seeks to examine what the trends in instructor evaluations are from the students at the University of Texas during the years 2000-2002. Among beauty, things such as age, gender, department, tenure, and how many students they were teaching are going to be taken into account and evaluated. Based on the observations from the 463 instructors, there is a correlation between the beauty of the instructor and the evaluation they have received.

>>>>

---
title: "How Beauty Trumps Performance: An Examination of Instructor Evaluations at the University of Texas"
author: "Marcin Jaczynski"
thanks: "Code and data are available at: https://github.com/jackchinski/instructor-ratings"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "At the end of each semester of school, students are able to rate their professors and teachers through course evaluations. However, there are concerns about the validity and reliability of these evaluations as a measure of teaching effectiveness. This paper examines the impact of instructor beauty on course evaluations at the University of Texas between 2000 and 2002, while also considering other factors that may influence ratings. The findings suggest that beauty does have a positive effect on evaluations, as such the use of the instructor evaluations for the use of hiring, firing, and promoting instructors remains controversial."
format: pdf
bibliography: references.bib
---

@@

Instructor ratings have a direct impact on the validity of an instructor or professor, and may be used for hiring, firing, promotion and consideration for tenure purposes. However, if beauty directly influences the instructor's evaluation, then such a metric is not a fair representation of the instructor, and therefore should be omitted in considerations for promotions, firing and tenure. While there are a lot of factors that play into how students will perceive an instructor and what kind of rating they will give them at the end of the semester, one that is not taken into consideration many times is how beautiful they are. As such, I have chosen to try and answer "How does an instructors beauty, according to their students, affect the instructors course evaluation?" This paper seeks to examine what the trends in instructor evaluations are from the students at the University of Texas during the years 2000-2002. Among beauty, things such as age, gender, department, tenure, and how many students they were teaching are going to be taken into account and evaluated. Based on the observations from the 463 instructors, there is a correlation between the beauty of the instructor and the evaluation they have received. These results have important implications for how we understand the factors that shape student perceptions of instructors and how we evaluate instructor performance. The correlation between beauty and evaluation is not strong, and does not persist when separating male and female instructors, where males average a higher evaluation yet score lower in terms of beauty than females. However, other things such as being native, and the class size also have an impact on the overall evaluation of the instructor. This paper will elaborate on the data, how it was sourced and analyzed, then will explore the results, and discuss them as well as their limitations.
